Thermodynamic response functions in a cell fluid model

### 2024-09-10 paper_text.tex
add reference to JML 2016; mention entropy as contribution to beta_v

Our modification contains an additional repulsive term, similarly to~\cite{MartinezValenciaEtAl2013}, as well as introduces parameter $\gamma$, which can vary as opposed to being strictly equal to 2 in the Morse potential. Including the repulsive term enables us to single out a reference system (in the reciprocal space) and apply the method of collective variables to calculating the grand partition function~[REFERENCE NEEDED].

@@

\subsection{Thermal pressure coefficient}
The thermal pressure coefficient is defined by
\begin{equation}
	\label{def:therm_pres_coef}
	\beta_V = \left( \frac{\partial P}{\partial T} \right)_{V,N}.
\end{equation}
We rewrite $\beta_V$ into a form that is suitable for the equation of state $P=P(T,\mu)$, see~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}
\begin{eqnarray}
	\beta_V	& = & \left(\frac{\partial P}{\partial T}\right)_{\mu} + \left(\frac{\partial P}{\partial \mu}\right)_T \left(\frac{\partial \mu}{\partial T}\right)_{V, N}.
\end{eqnarray}
Applying the cyclic relation
\begin{equation*}
	\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
	\left(\frac{\partial T}{\partial V}\right)_{\mu, N}
	\left(\frac{\partial V}{\partial \mu}\right)_{T, N}
	= -1
\end{equation*}
we obtain
\begin{eqnarray*}
	\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
	& = & 
	- \left(\frac{\partial V}{\partial T}\right)_{\mu, N} 
	\left(\frac{\partial V}{\partial \mu}\right)^{-1}_{T, N} \\
	& = & - \left(\frac{\partial \bar{n}}{\partial T}\right)_{\mu}
	\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T}.		
\end{eqnarray*}
Substituting this result into the formula for $\beta_V$, we arrive at the final expression for the thermal pressure coefficient
\begin{equation}
	\beta_V = \left(\frac{\partial P}{\partial T}\right)_{\mu} 
	- \left(\frac{\partial P}{\partial \mu}\right)_T 
	\left(\frac{\partial \bar{n}}{\partial T}\right)_{\mu}
	\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T}
\end{equation}
which\footnote{Compare this equation for $\beta_V$ with Eq.(17) from~\cite{StrokerMeier2021}} is easily calculated based on Eqs.~\eqref{eq:eosMT} and~\eqref{eq:density}.

>>>>

Our modification contains an additional repulsive term, similarly to~\cite{MartinezValenciaEtAl2013}, as well as introduces parameter $\gamma$, which can vary as opposed to being strictly equal to 2 in the Morse potential. Including the repulsive term enables us to single out a reference system (in the reciprocal space) and apply the method of collective variables to calculating the grand partition function~\cite{KozlovskiiDobush2016}.

@@

\subsection{Thermal pressure coefficient}
The thermal pressure coefficient is defined by
\begin{equation}
	\label{def:therm_pres_coef}
	\beta_V = \left( \frac{\partial P}{\partial T} \right)_{V,N}.
\end{equation}
We rewrite $\beta_V$ into a form that is suitable for the equation of state $P=P(T,\mu)$, see~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}
\begin{eqnarray}
	\beta_V	& = & \left(\frac{\partial P}{\partial T}\right)_{\mu} + \left(\frac{\partial P}{\partial \mu}\right)_T \left(\frac{\partial \mu}{\partial T}\right)_{V, N}.
\end{eqnarray}
Applying the cyclic relation
\begin{equation*}
	\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
	\left(\frac{\partial T}{\partial V}\right)_{\mu, N}
	\left(\frac{\partial V}{\partial \mu}\right)_{T, N}
	= -1
\end{equation*}
we obtain
\begin{eqnarray*}
	\left(\frac{\partial \mu}{\partial T}\right)_{V, N} 
	& = & 
	- \left(\frac{\partial V}{\partial T}\right)_{\mu, N} 
	\left(\frac{\partial V}{\partial \mu}\right)^{-1}_{T, N} \\
	& = & - \left(\frac{\partial \bar{n}}{\partial T}\right)_{\mu}
	\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T}.		
\end{eqnarray*}
Substituting this result into the formula for $\beta_V$, we arrive at the final expression for the thermal pressure coefficient
\begin{equation}
	\beta_V = \left(\frac{\partial P}{\partial T}\right)_{\mu} 
	- \left(\frac{\partial P}{\partial \mu}\right)_T 
	\left(\frac{\partial \bar{n}}{\partial T}\right)_{\mu}
	\left(\frac{\partial \bar{n}}{\partial \mu}\right)^{-1}_{T}
\end{equation}
which\footnote{Compare this equation for $\beta_V$ with Eq.(17) from~\cite{StrokerMeier2021}} is easily calculated based on Eqs.~\eqref{eq:eosMT} and~\eqref{eq:density}. It is also worth noting that the first contribution to $\beta_V$ is essentially the entropy per volume, $S/V = (\partial P / \partial T)_\mu$.
